Brief Announcement: An Improved Lower Bound for Dynamic Balanced Graph Partitioning

### 2020-02-14 paper/paper.tex
Merge by hand

  We say that \TAlg{} is \emph{component respecting}, i.e., it maintaints the invariant that nodes of each component are located at the same server.

  \begin{lemma}
    \label{lem:1req}
    Serving each request incurs the cost at most $3\cdot\alpha$ for \TAlg.
  \end{lemma}

  \begin{proof}
    If no partition of \TAlg's components exist, then it performs no repartition, and it pays nothing.
    Hence, in the following we bound the cost of moving to component-respecting partition.

@@

  \begin{theorem}
    \TAlg{} is $O(\ell)$-competitive.
  \end{theorem}
  \begin{proof}
    Consider a request $(u,v)$.
    If $u$ and $v$ are collocated in the same server, then \TAlg{} pays nothing.

    \TAlg{} collocates nodes belonging to one component, hence during a single phase, we have at most $2\cdot \ell$ requests to non-collocated nodes.
    We use Lemma~\ref{lem:1req} to bound the cost of each such request, and therefore \TAlg{} pays $6\cdot\alpha\cdot \ell$ in a single phase.
    On the other hand, during each phase, OPT pays at least $1$.
  \end{proof}

>>>>

  We say that \TAlg is \emph{component respecting}, i.e., it maintains the invariant that nodes of each component are located at the same server.
  Observe that this implies that \TAlg never increases any counter above $\alpha$.

  \begin{lemma}
    \label{lem:1req}
    A single repartition incurs the cost at most $3\cdot\alpha$ for \TAlg.
  \end{lemma}

  \begin{proof}
    If no partition of \TAlg's components exist, then it performs no repartition, and serves the request remotely, incurring the cost $1$.
    Hence, in the following we bound the cost of moving to component-respecting partition.

@@

  \begin{theorem}
    \TAlg{} is $O(\ell)$-competitive.
  \end{theorem}
  \begin{proof}
    Fix a phase of \TAlg, and consider the state of its counters at the end of it.

    Let $A^I$ be the cost of (extra-cluster) communication incurred in this phase by \TAlg between pairs that belonged to the single component at the end of the phase.
    Let $A^E$ be the cost of (extra-cluster) communication incurred in this phase between the nodes that belong to different components throughout this phase.
    Let $A^M$ be the cost of migrations performed by \TAlg in this phase.

    We say that the pair $(u, v)$ is \emph{saturated}, if the counter has value $\alpha$, and \emph{unsaturated} otherwise.

    During a single phase, at most $2\cdot \ell$ component merge operations are performed.
    Exceeding this number would mean that a component of size $4$ would emerge, a contradiction.
    Together with Lemma~\ref{lem:1req}, this gives us $A^M \leq 6\cdot\alpha\cdot\ell$.

    Now we bound $A^I$.
    Each component of size $3$ contributes at most $(3\cdot \alpha - 1)$ to $A^I$, as at most $2$ of its pairs are saturated and contribute $\alpha$ each, and the unsaturated pair contributes at most $\alpha-1$.
    Each of $\ell$ clusters of \TAlg contain at most one component of size $3$ (other cluster configurations contribute less: $3$ singletons contribute $0$ and a component of size $2$ contributes at most $\alpha$ to $A^I$), and in total, $A^I \leq \ell \cdot (3\cdot \alpha-1)$.

    We bound $A^E$ by $k^2 \cdot (\alpha - 1)$, as no more than $k^2$ pairs are unsaturated, and each of them contributes at most $\alpha -1$.

    In total, the cost of \TAlg is at most $A^I + A^E + A^M$ during the phase.

    Now we lower bound the cost of $\OPT$.
  \end{proof}
